Learned image compression

### 2023-03-14 paper/paper.tex
conclusions

The model was trained with multiple different configurations. We tried $\lambda \in \{0.0008, 0,001\}$ where latent space channels $M$ and $N$ were $M=N=192$ and $\lambda \in \{0.0004, 0.0008\}$ where latent space channels $M$ and $N$ were $M=192, N=128$.
All of these models were trained for approximately 300 epochs but they could be still trained further.
Unfortunately due to lack of computational capabilities we could not train more networks and even these networks most likely could have been trained for longer but with epochs taking from 180 seconds (for smaller latent space) to 300 seconds (larger latent space), we could not train them further due to time constraints.

>>>>

The model was trained with multiple different configurations. We tried $\lambda \in \{0.0008, 0.001\}$ where latent space channels $M$ and $N$ were $M=N=192$ and $\lambda \in \{0.0004, 0.0008\}$ where latent space channels $M$ and $N$ were $M=192, N=128$.
All of these models were trained for approximately 300 epochs but they could be still trained further.
Unfortunately due to lack of computational capabilities we could not train more networks and even these networks most likely could have been trained for longer but with epochs taking from 180 seconds (for smaller latent space) to 300 seconds (larger latent space), we could not train them further due to time constraints.

\section{Discussions and Conclusions}
The goal of this work was to implement and verify the methods of He \textit{et al.} 2022 \cite{ELIC} for learned image compression. We successfully implemented and trained 4 different models targeted to low compression ratios by varying the model hyperparameters $M$, $N$ and $\lambda$ in loss function. The implementation differs from the paper in few details as discussed earlier. Due to limited time and computing resources we were not able to train our models to convergence and we used fewer images in our training set. Finetuning of the models using MS-SSIM as loss function was also discarded due to computation and time restrictions.

The results from our models were far from the level of original ELIC paper in terms of subjective and objective measurements in quality. This was caused by implementation differences and different training setups. Additionally, our models suffer from color artifacts that were introduced in various areas in result images. We are not sure what is the cause of these artifacts, but possible reasons are errors in implementation for example overflow in data types or underfitted models. In subjective evaluation we were able to pass the level of JPEG compression, but objective metrics were orders of magnitude behind JPEG. We can conclude that were where able to have working image compression models using methods introduced by He \textit{et al.} 2022 \cite{ELIC} in ELIC, but we can not confirm the improved quality and inference speed they achieved. We neither studied the progressive decoding or the fast thumbnail generator introduced in original paper.

Apparently the models from original ELIC paper also suffers from some kind of artifacts that they do not mention in paper. In successive work He \textit{et al.} 2022 \cite{PO-ELIC} propose a discriminator network in top of ELIC to reduce amount of know artifacts in learned image compression with Generative adversarial network (GAN) inspired training. Other interesting directions in learned image compression are purely GAN and diffusion based compressors, that are capable of reconstructing even unseen areas of image using techniques of generative methods \cite{HIFIC, diffusion}.
